---
title: "Drawing DAGs and SWIGs for Mendelian randomization analyses"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Drawing DAGs and SWIGs for Mendelian randomization analyses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(out.width = "65%")
```

# Introduction

This webpage shows some R code for generating figures, directed acyclic graphs (DAGs), and single world intervention graphs (SWIGs) for Mendelian randomization (MR) analyses.

First, we load in the tidyverse collection of packages to have access to the pipe and the ggplot2 package, amongst other features.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

If you do not already have one you will need a LaTeX installation. In R it is easiest to install [TinyTeX](https://yihui.org/tinytex/) as follows.
```{r, eval=FALSE}
install.packages("tinytex")
tinytex::install_tinytex()
```

# DiagrammeR package

The website for DiagrammeR is here: http://rich-iannone.github.io/DiagrammeR/ .

We will draw the DAGs using Graphviz which uses the DOT language to describe graphs.

## Single genotype DAG

* Version without borders around observed variables

```{r}
mrdags::single_iv_dag()
```

* Version with borders around observed variables

```{r}
mrdags::single_iv_dag(noborders = FALSE)
```

## Multiple genotype DAG

* Version without borders around observed variables

```{r}
mrdags::multiple_iv_dag()
```

* Version with borders around observed variables

```{r}
mrdags::multiple_iv_dag(noborders = FALSE)
```

This could be extended for more genotypes.

## Single genotype path diagram

Explicitly including the confounder, $U$.
```{r}
mrdags::single_iv_path_diagram()
```

Instead drawing correlated error terms between $X$ and $Y$.

```{r}
mrdags::single_iv_path_diagram(confounder = FALSE)
```

# dagitty package

The website for DAGitty is here http://www.dagitty.net/

```{r, warning=FALSE, message=FALSE}
library(dagitty)
```

## Using DOT syntax

The DAG for a single genotype.
```{r, fig.height=5, fig.width=7}
mrdag <- dagitty('dag {
  G -> X -> Y
  X <- U -> Y
  G [pos="0,0"]
  X [e, pos="1,0"]
  U [pos="1.5,-0.4"]
  Y [o, pos="2,0"]
}')

plot(mrdag)
```

## Using ggdag

```{r, fig.height=5, fig.width=7}
mrdags::ggdag_single_iv_dag()
```

This can be labelled as follows.

```{r, message=FALSE, warning=FALSE}
library(ggdag)
```

```{r, fig.height=5, fig.width=7}
mrdag <- mrdag %>%
  dag_label(
    labels = c(
      "X" = "Exposure",
      "Y" = "Outcome",
      "G" = "Genotype",
      "U" = "Confounders"))
ggdag(mrdag,  use_labels = "label") + theme_dag()
```

Dagitty can identify the instrumental variable on the DAG.

```{r, fig.height=5, fig.width=7}
mrdag %>% ggdag_instrumental() + theme_dag()
```

It correctly identifies the genotype, G, as the instrumental variable.

## Using R model type syntax in ggdag

```{r, fig.height=5, fig.width=7}
coords <- list(x = c(
  G = 0,
  X = 1,
  Y = 2,
  U = 1.5
),
y = c(
  G = 0,
  X = 0,
  Y = 0,
  U = 1
))

mrdag <- dagify(
  X ~ G + U,
  Y ~ X + U,
  labels = c(
    "G" = "Genotype",
    "X" = "Exposure",
    "Y" = "Outcome",
    "U" = "Confounders"
  ),
  exposure = "X",
  outcome = "Y",
  coords = coords
)

ggdag(mrdag, use_labels = "label") + theme_dag()
```

Plotting directly with ggplot2 functions.

```{r, fig.height=5, fig.width=7}
mrdag %>%
  ggplot(aes(
    x = x,
    y = y,
    xend = xend,
    yend = yend
  )) +
  geom_dag_point() +
  geom_dag_edges() +
  geom_dag_text() +
  theme_dag()
```

Showing the paths.

```{r, fig.height=5, fig.width=7}
dagify(Y ~ X + U,
       X ~ U + G,
       exposure = "X",
       outcome = "Y") %>%
ggdag_paths() + theme_dag()
```

Adjustment set for the effect of $X$ on $Y$.

```{r, fig.height=5, fig.width=7}
dagify(Y ~ X + U,
       X ~ U + G,
       exposure = "X",
       outcome = "Y",
       coords = coords) %>%
  ggdag_adjustment_set() + theme_dag()
```

This shows that the path from G to X is unconfounded. And that when U is adjusted for the path from X to Y is unconfounded.

## Multiple genotype DAG

```{r, fig.height=5, fig.width=7}
mrdags::ggdag_multiple_iv_dag()
```

This could be extended for more genotypes.

# Eleanor Murray's TikZ diagrams using the tikz code chunk engine

I believe that tikz diagrams can be included in Rmarkdown documents using the tikz code chunk engine for output types `pdf_document` and `html_document`.

Tikz is a package for LaTeX and so pdf is its natural output. So far I have not been able to include these figures in either `html_notebook` or `word_document` output formats. I suspect that the images need conversion to png files for these to work under rmarkdown.

We will use some of the LaTeX code for drawing DAGs using tikz provided by Eleanor Murray [here](https://github.com/eleanormurray/causalgraphs_latex).

In its LaTeX preamble, this code loads in several tikz libraries. As such for our tikz code chunks to compile in our Rmd file we need to define a new template file, e.g.`tikz-template.tex`, which is provided in the **mrdags** package.

If you have a new TinyTeX install check you have the following LaTeX packages installed.

```{r, eval=FALSE}
tinytex::tlmgr_install(c("preview", "pgf"))
```

We can now plot our DAGs.

## Single genotype DAG using tikz

We call this in the following way in our R Markdown code

````{verbatim}
```{r, child=mrdags::tikz_single_iv_dag()}
```
````

which produces the following.

```{r, child=mrdags::tikz_single_iv_dag()}
```

## Multiple genotype DAG using tikz

We call this in the following way in our R Markdown code

````{verbatim}
```{r, child=mrdags::tikz_multiple_iv_dag()}
```
````

which produces the following.

```{r, child=mrdags::tikz_multiple_iv_dag()}
```

Again this could be extended for more genotypes.

# quickdag package

We can produce some nice plots with the quickdag package, which appears to be based on Eleanor Murray's TikZ code, from [here](https://github.com/jrgant/quickdag)

```{r, warning=FALSE}
if (!requireNamespace("quickdag", quietly = TRUE)) remotes::install_github("jrgant/quickdag")

library(quickdag)

edges <- c("G -> X",
           "X -> Y",
           "U -> {X Y}")

# DAG
dag  <- qd_dag(edges, verbose = FALSE)
dag %>% render_graph()
```

We can also draw the single world intervention graph (SWIG) for the IV model as follow.
```{r}
# SWIG
swig <- dag %>%
  qd_swig(fixed.nodes = c("G","X"),
          custom.labels = c("G" = "g",
                            "X" = "x"))
swig %>% render_graph()
```

# Thomas Richardson's TikZ code for SWIGs

Thomas Richardson provides a TikZ/PGF shape library for SWIGs called tikz-swigs, which is available on CTAN [here](https://ctan.org/pkg/tikz-swigs).

* Install the package into your TinyTeX installation with
    ```{r, eval=FALSE}
    tinytex::tlmgr_install(c("preview", "pgf", "tikz-swigs"))
    ```
* Its user manual is available [here](http://mirrors.ctan.org/graphics/pgf/contrib/tikz-swigs/tikz-swigs.pdf)

To use the tikz chunk engine we again write a LaTeX template based on the preamble in the article by Richardson. The LaTeX template is included in the package.

We will recreate the SWIG template in Figure 2 (b) of Swanson et al. [here](https://doi.org/10.1080/01621459.2018.1434530).

```{r, child=mrdags::tikz_swig_iv_bw()}
```

Let's add colour in the same way as in the original figure.

```{r, child=mrdags::tikz_swig_iv_color()}
```

## SWIGs for MVMR models

Let's attempt the SWIG for the multivariable Mendelian randomization (MVMR) model.

### SWIG for MVMR model with 2 independent exposures

Since $G_1$ an instrument for both $X_1$ and $X_2$, $X_2$ represents a pleiotropic effect in this case.

```{r, child=mrdags::tikz_swig_mvmr_pleiotropy()}
```

### SWIG for MVMR model with $X_2$ confounder of $X_1$--$Y$

```{r, child=mrdags::tikz_swig_mvmr_confounder()}
```

### SWIG for MVMR model with $X_2$ collider of $X_1$--$Y$

```{r, child=mrdags::tikz_swig_mvmr_collider()}
```

### SWIG for MVMR model with $X_2$ mediator of $X_1$--$Y$

```{r, child=mrdags::tikz_swig_mvmr_mediator()}
```
