---
title: "Code to plot DAGs for Mendelian randomization analyses"
author: "Tom Palmer"
date: '2019-01-24'
output:
  html_notebook:
    toc: yes
    toc_depth: 2
always_allow_html: yes
---

This webpage shows the code for generating figures for Mendelian randomization analyses.

First we load in the tidyverse collection of packages to have access to the pipe and the ggplot2 package, amongst other features.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

# DiagrammeR package

The website for DiagrammeR is here: http://rich-iannone.github.io/DiagrammeR/ .

```{r, warning=FALSE}
library(DiagrammeR)
```

## Single genotype

```{r}
grViz("
      digraph mrdag {

      graph [rankdir=TB]

      node [shape=ellipse]
      U [label='Confounders']

      node [shape=box]
      G [label='Genotype']
      X [label='Phenotype']
      Y [label='Outcome']

      { rank = same; G X Y }

      G -> X [minlen=3]
      U -> X
      U -> Y
      X -> Y [minlen=3]
      }
      ")
```

```{r, include=FALSE}
#library(DiagrammeRsvg)
#svg <- export_svg(grvizdag)
```

## Multiple genotypes

```{r}
grViz("
      digraph mrdag {

      graph [rankdir=LR]

      node [shape=ellipse]
      U [label='Confounders']

      node [shape=box]
      G1 [label='Genotype 1']
      G2 [label='Genotype 2']
      G3 [label='Genotype 3']
      X [label='Phenotype']
      Y [label='Outcome']

      { rank = same; G1 G2 G3 }

      G1 -> X 
      G2 -> X 
      G3 -> X 
      U -> X
      U -> Y
      X -> Y
      }
      ")
```

# dagitty package

The website for DAGitty is here http://www.dagitty.net/ .

```{r, warning=FALSE}
library(dagitty)
```

## Using dot syntax

```{r}
mrdag <- dagitty('dag {
  G -> X -> Y
  X <- U -> Y
  G [pos="0,0"]
  X [e, pos="1,0"]
  U [pos="1.5,-1"]
  Y [o, pos="2,0"]
}')

plot(mrdag)
```

## Version using ggdag

```{r, warning=FALSE}
library(ggdag)
```

```{r}
mrdag <- dagitty('dag {
  G -> X -> Y
  X <- U -> Y
  G [pos="0,0"]
  X [e, pos="1,0"]
  U [pos="1.5,1"]
  Y [o, pos="2,0"]
}')

ggmrdag <- tidy_dagitty(mrdag)
ggdag(ggmrdag)
```

Alternatively we could position $U$ over $X$.
```{r}
mrdag <- dagitty('dag {
  G -> X -> Y
  X <- U -> Y
  G [pos="0,0"]
  X [e, pos="1,0"]
  U [pos="1,1"]
  Y [o, pos="2,0"]
}')

ggmrdag <- tidy_dagitty(mrdag)
ggdag(ggmrdag)
```

This can be labelled as follows.
```{r}
mrdag <- mrdag %>% 
  dag_label(
    labels = c(
      "X" = "Exposure", 
      "Y" = "Outcome", 
      "G" = "Genotype",
      "U" = "Confounders"))
ggdag(mrdag,  use_labels = "label")
```

Dagitty can identify the instrumental variable on the DAG.
```{r}
mrdag %>% ggdag_instrumental()
```

## Using ggdag dagify R model type syntax

```{r}
coords <- list(x = c(
  G = 0,
  X = 1,
  Y = 2,
  U = 1.5
),
y = c(
  G = 0,
  X = 0,
  Y = 0,
  U = 1
))

mrdag <- dagify(
  X ~ G + U,
  Y ~ X + U,
  labels = c(
    "G" = "Genotype",
    "X" = "Exposure",
    "Y" = "Outcome",
    "U" = "Confounders"
  ),
  exposure = "X",
  outcome = "Y",
  coords = coords
)

ggdag(mrdag, use_labels = "label")
```

Plotting directly with ggplot2 functions.
```{r}
mrdag %>%
  ggplot(aes(
    x = x,
    y = y,
    xend = xend,
    yend = yend
  )) +
  geom_dag_point() +
  geom_dag_edges() +
  geom_dag_text() +
  theme_dag()
```

Showing the paths.
```{r}
dagify(Y ~ X + U,
       X ~ U + G,
       exposure = "X",
       outcome = "Y") %>%
ggdag_paths()
```

Adjustment set for the effect of $X$ on $Y$.  
```{r}  
dagify(Y ~ X + U,
       X ~ U + G,
       exposure = "X",
       outcome = "Y", 
       coords = coords) %>%
  ggdag_adjustment_set()
```  

## Multiple genotypes

```{r}
coords2 <- list(x = c(
  G1 = 0,
  G2 = 0,
  G3 = 0,
  X = 1,
  Y = 2,
  U = 1.5
),
y = c(
  G1 = 0.5,
  G2 = 0,
  G3 = -0.5,
  X = 0,
  Y = 0,
  U = 1
))


mrdag <- dagify(
  X ~ G1,
  X ~ G2,
  X ~ G3,
  Y ~ X,
  X ~ U,
  Y ~ U,
  exposure = "X",
  outcome = "Y",
  coords = coords2
)

ggdag(mrdag)
```
